---
title: "Mixture models"
output: 
  html_document:
    toc: true
date: "2024-03-14"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## FCS file import


```{r}
library(flowCore)
```


```{r}
file.name <- system.file("extdata","0877408774.B08", 
                         package="flowCore")
x <- read.FCS(file.name, transformation=FALSE)
summary(x)
```


## MASS::fitdistr


The package 'MASS' is a part of R, so it does not require installation.
This may make it appear to be a first option.
The function MASS::fitdistr() will fit a **single** distribution of many sorts.
This makes it a great option if the goal is to fit a single distribution.


```{r}
exprs(x)[1:3,]
# plot(exprs(x)[, 1])

hist(exprs(x)[, "FL1-H"], breaks = seq(0, 1e3, by = 10), border = NA)
my_params <- MASS::fitdistr(exprs(x)[, "FL1-H"], densfun = "normal")

hist(exprs(x)[, "FL1-H"], freq = FALSE, breaks = seq(0, 1e3, by = 10), border = NA)
my_x <- seq(0, 1e3, by=1)
my_y <- dnorm( my_x, mean = my_params$estimate[1], sd = my_params$estimate[2])
lines(x = my_x, y = my_y)
```


The data appears to be at least bimodal in distribution.
A peak a bit above 200, another at just over 600.
There may also be a peak near zero.
The peak near zero may not be of research interest, but may be worth accounting for.
For example, a peak near zero may represent errors that should be accounted for so that they can be separated from expected results.


## Mixture models on CRAN


Mixture models are methodas that fit more than one distribution to the data.
This may be multiple distributions of the same form.
For example, a mixture model may fit 3 distributions, but they are are **normal distributions**.
Some methods may allow for multiple distributions that include different forms of distribution.
For example, a mixture model may fit one uniform distribution, 1 Poisson distribution, and 1 normal distributions.
Different methods provide different options, so some exploration may be required to find an implementation with the options desired for a specific application.


In the R language, a good starting point when looking for a new method is at CRAN.
One strategy I use when looking for a new method is to use the CRAn packages sorted by name page

https://cran.r-project.org/web/packages/available_packages_by_name.html

and use the 'Find' method in the browser to query the page for a keyword.
On 2024-03-15 I queried this page for 'mixture model' and found the following.

- AdaptGauss
- aldvmm
- BAMBI
- batchmix
- bayesanova    
- BNPmix    
- flexmix	Flexible Mixture Modeling    
- mclust	Gaussian Mixture Modelling for Model-Based Clustering, Classification, and Density Estimation
- MixSemiRob	Mixture Models: Parametric, Semiparametric, and Robust
- mixtools	Tools for Analyzing Finite Mixture Models
- mixture	Mixture Models for Clustering and Classification

These are packages that seemed to stick out to me while I was browsing the 73 matches.


## mixtools


https://cran.r-project.org/web/packages/mixtools/


```{r}
my_dat <- exprs(x)
my_dat <- as.data.frame(my_dat)
my_dat[1:3, ]

library("mixtools")

mm1 <- normalmixEM(my_dat$`FL1-H`, lambda = .5, mu = c(200, 600), sigma = c(60, 70))

# plot(mm1, density = TRUE, cex.axis = 1.4, cex.lab = 1.4, cex.main = 1.8,
#      main2 = "Time between Old Faithful eruptions", xlab2 = "Minutes")

mm1[c("lambda", "mu", "sigma")]

#summary(mm1)

hist(exprs(x)[, "FL1-H"], freq = FALSE, breaks = seq(0, 1e3, by = 10), border = NA)

my_x <- seq(0, 1e3, by=1)
my_y1 <- dnorm( my_x, mean = mm1$mu[1], sd = mm1$sigma[1] ) * mm1$lambda[1]
my_y2 <- dnorm( my_x, mean = mm1$mu[2], sd = mm1$sigma[2] ) * mm1$lambda[2]

lines(x = my_x, y = my_y1)
lines(x = my_x, y = my_y2)

```


```{r}
mm1 <- normalmixEM(my_dat$`FL1-H`, lambda = .5, mu = c(50, 200, 600), sigma = c(100, 50, 50))
mm1[c("lambda", "mu", "sigma")]

#summary(mm1)

hist(exprs(x)[, "FL1-H"], freq = FALSE, breaks = seq(0, 1e3, by = 10), border = NA)

my_x <- seq(0, 1e3, by=1)
my_y1 <- dnorm( my_x, mean = mm1$mu[1], sd = mm1$sigma[1] ) * mm1$lambda[1]
my_y2 <- dnorm( my_x, mean = mm1$mu[2], sd = mm1$sigma[2] ) * mm1$lambda[2]
my_y3 <- dnorm( my_x, mean = mm1$mu[3], sd = mm1$sigma[2] ) * mm1$lambda[3]
lines(x = my_x, y = my_y1)
lines(x = my_x, y = my_y2)
lines(x = my_x, y = my_y3)

```


```{r}
#dat_mult <- makemultdata()
#mm1 <- multmixEM(my_dat$`FL1-H`, lambda = 5)


```




## flexmix


https://cran.r-project.org/web/packages/flexmix/


```{r}
# library("flexmix")

```




