---
title: "OrthoFinder"
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: hide
date: "2023-09-21"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.align = "center")
```


## Links

- https://github.com/davidemms/OrthoFinder
- https://davidemms.github.io/orthofinder_tutorials/exploring-orthofinders-results.html


## Species tree


The OrthoFinder species tree describes the relationships among each sample.
The species tree includes node labels which are used in other output as well.
For example, the Phylogenetic Hierarchical Orthogroups are organized into files based on nodes of the species tree (N0, N1, etc.).
This species tree can be read into R and visualized using ape and/or ggtree.


### ape


```{r}
library(ape)
my_tree <- read.tree("../Species_Tree/SpeciesTree_rooted_node_labels.txt")
# my_tree
# my_tree$node.label
my_tree$tip.label

plot(my_tree)
nodelabels( text = my_tree$node.label, frame = "none", bg = "#FFFFFF", adj = c(1.1, -0.1))
axis(side = 1)
```


### ggtree


```{r}
suppressMessages(library(ggtree))
library(ggplot2)

#ggtree(my_tree)
p <- ggtree(my_tree) + geom_tiplab() 
# p
p <- p + theme_tree2(plot.margin=margin(16, 30, 16, 30)) +
    scale_x_continuous( limits = c(0, 0.4), breaks = seq(0, 0.4, by = 0.1),
                        labels = seq(0, 0.4, by = 0.1) )
# p

# p + geom_label( aes(x=branch, label = 1:31 ), fill='lightgreen')
#p + geom_text(aes(label=c(my_tree$tip.label, my_tree$node.label)), hjust=1.1)
p <- p + geom_text(aes(label=c(rep("", times = length(my_tree$tip.label)), my_tree$node.label)), hjust=1.1, vjust=-0.8)
p
```



## Phylogenetic_Hierarchical_Orthogroups

The OrthoFinder output includes a directory of Phylogenetic Hierarchical Orthogroups.
This is a directory of files named by the clades in the species tree.
The clade describes which samples have genes in the group and which do not.


```{r}

pho <- read.table(file = "../Phylogenetic_Hierarchical_Orthogroups/N0.tsv", sep = "\t", header = TRUE)
pho[c(1:2, 6:10), 1:5]
names(pho)
nrow(pho)

par(mar = c(10, 4, 4, 2))
barplot(apply(pho[, -c(1:3)], MARGIN = 2, function(x){ sum(!is.na(x)) }), las = 3)
title( main = "N0" )
par(mar = c(5, 4, 4, 2))

pho <- read.table(file = "../Phylogenetic_Hierarchical_Orthogroups/N1.tsv", sep = "\t", header = TRUE)
pho[c(1:2, 6:10), 1:5]
names(pho)
nrow(pho)

par(mar = c(10, 4, 4, 2))
barplot(apply(pho[, -c(1:3)], MARGIN = 2, function(x){ sum(!is.na(x)) }), las = 3)
title( main = "N1" )
par(mar = c(5, 4, 4, 2))
#barplot(apply(pho[, -c(1:3)], MARGIN = 2, function(x){ sum(is.na(x)) }), las = 3)

pho <- read.table(file = "../Phylogenetic_Hierarchical_Orthogroups/N2.tsv", sep = "\t", header = TRUE)
pho[c(1:2, 6:10), 1:5]
names(pho)
nrow(pho)

par(mar = c(10, 4, 4, 2))
barplot(apply(pho[, -c(1:3)], MARGIN = 2, function(x){ sum(!is.na(x)) }), las = 3)
title( main = "N2" )
par(mar = c(5, 4, 4, 2))

pho <- read.table(file = "../Phylogenetic_Hierarchical_Orthogroups/N3.tsv", sep = "\t", header = TRUE)
pho[c(1:2, 6:10), 1:5]
names(pho)
nrow(pho)

par(mar = c(10, 4, 4, 2))
barplot(apply(pho[, -c(1:3)], MARGIN = 2, function(x){ sum(!is.na(x)) }), las = 3)
title( main = "N3" )
par(mar = c(5, 4, 4, 2))


```


