---
title: "WGCNA tutorial"
author: "Insert authors here"
date: "`r format(Sys.time(), '%Y, %B %d')`"
output:
  html_document:
    toc: true
    code_folding: show
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.align = "center")
```

# Background

Tutorial for the WGCNA CRAN package (https://CRAN.R-project.org/package=WGCNA) adapted from 'Tutorials for the WGCNA package' (https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/index.html).


# Simulation of expression and trait data


```{r}
library(WGCNA)
```


```{r}
# Here are input parameters of the simulation model
# number of samples or microarrays in the training data
no.obs=50
# now we specify the true measures of eigengene significance
# recall that ESturquoise=cor(y,MEturquoise)
ESturquoise=0;   ESbrown= -.6;
ESgreen=.6;ESyellow=0
# Note that we don’t specify the eigengene significance of the blue module
# since it is highly correlated with the turquoise module.
ESvector=c(ESturquoise,ESbrown,ESgreen,ESyellow)
# number of genes 
nGenes1=3000
# proportion of genes in the turquoise, blue, brown, green, and yellow module #respectively.
simulateProportions1=c(0.2,0.15, 0.08, 0.06, 0.04)
# Note that the proportions don’t add up to 1. The remaining genes will be colored grey,
# ie the grey genes are non-module genes.
# set the seed of the random number generator. As a homework exercise change this seed.
set.seed(1)
#Step 1: simulate a module eigengene network.
# Training Data Set I
MEgreen=rnorm(no.obs)
scaledy=MEgreen*ESgreen+sqrt(1-ESgreen^2)*rnorm(no.obs)
y=ifelse( scaledy>median(scaledy),2,1)
MEturquoise= ESturquoise*scaledy+sqrt(1-ESturquoise^2)*rnorm(no.obs)
# we simulate a strong dependence between MEblue and MEturquoise
MEblue= .6*MEturquoise+ sqrt(1-.6^2) *rnorm(no.obs)
MEbrown= ESbrown*scaledy+sqrt(1-ESbrown^2)*rnorm(no.obs)
MEyellow= ESyellow*scaledy+sqrt(1-ESyellow^2)*rnorm(no.obs)
ModuleEigengeneNetwork1=data.frame(y,MEturquoise,MEblue,MEbrown,MEgreen, MEyellow)
```


```{r}
ModuleEigengeneNetwork1[1:3, ]
nrow(ModuleEigengeneNetwork1)
```


Simulate data.


```{r chunk 1.3}
dat1=simulateDatExpr5Modules(MEturquoise=ModuleEigengeneNetwork1$MEturquoise,
          MEblue=ModuleEigengeneNetwork1$MEblue,
          MEbrown=ModuleEigengeneNetwork1$MEbrown,
          MEyellow=ModuleEigengeneNetwork1$MEyellow,
          MEgreen=ModuleEigengeneNetwork1$MEgreen, 
          nGenes=nGenes1, 
          simulateProportions=simulateProportions1)

class(dat1)
str(dat1)

dat1$datME[1:3, ]
dat1$datExpr[1:5, 1:3]
```


```{r chunk_1.4}
datExpr = dat1$datExpr
truemodule = dat1$truemodule
datME = dat1$datME
# attach(ModuleEigengeneNetwork1)
```


```{r chunk_1.5}
table(truemodule)
dim(datExpr)
```

`r ncol(datExpr)` genes (rows) and `r nrow(datExpr)` samples (columns).
'Arrays' are samples (columns).


```{r chunk_1.6}
datExpr=data.frame(datExpr)
ArrayName=paste("Sample",1:dim(datExpr)[[1]], sep="" )   
# The following code is useful for outputting the simulated data 
GeneName=paste("Gene",1:dim(datExpr)[[2]], sep="" )   
dimnames(datExpr)[[1]]=ArrayName
dimnames(datExpr)[[2]]=GeneName
```



# Loading of expression data, an alternative to data simulation, provided to illustrate data loading of real data


```{r chunk_2.1}
# Load WGCNA package
library(WGCNA)
```


```{r chunk_2.2, eval=FALSE}
datGeneSummary=read.csv("GeneSummaryTutorial.csv")
datTraits=read.csv("TraitsTutorial.csv")
datMicroarrays=read.csv("MicroarrayDataTutorial.csv")
```


```{r chunk_2.3, eval=FALSE}
#=====================================================================================
#
#  Code chunk 3
#
#=====================================================================================

# This vector contains the microarray sample names
ArrayName= names(data.frame(datMicroarrays[,-1]))
 # This vector contains the gene names
GeneName= datMicroarrays$GeneName
# We transpose the data so that the rows correspond to samples and the columns correspond to genes
# Since the first column contains the gene names, we exclude it.
datExpr=data.frame(t(datMicroarrays[,-1]))
names(datExpr)=datMicroarrays[,1]
dimnames(datExpr)[[1]]=names(data.frame(datMicroarrays[,-1]))
#Also, since we simulated the data, we know the true module color:
truemodule= datGeneSummary$truemodule
rm(datMicroarrays)
collectGarbage()


#=====================================================================================
#
#  Code chunk 4
#
#=====================================================================================


# First, make sure that the array names in the file datTraits line up with those in the microarray data 
table( dimnames(datExpr)[[1]]==datTraits$ArrayName)
# Next, define the microarray sample trait 
y=datTraits$y
```




# Basic data preprocessing illustrates rudimentary techniques for handling missing data and removing outliers

```{r chunk_3.1}
# Load WGCNA package
library(WGCNA)
# The following setting is important, do not omit.
#options(stringsAsFactors = FALSE);
# Load the previously saved data
# load("Simulated-dataSimulation.RData");
# attach(ModuleEigengeneNetwork1)
```


```{r chunk_3.2}
datExpr[1:3, 1:6]

meanExpressionByArray=apply( datExpr,1,mean, na.rm=T)  
NumberMissingByArray=apply( is.na(data.frame(datExpr)),1, sum)
```


```{r chunk_3.3}
# sizeGrWindow(9, 5)
barplot(meanExpressionByArray,
        xlab = "Sample", ylab = "Mean expression",
        main ="Mean expression across samples",
        names.arg = c(1:50), cex.names = 0.7)
```


```{r chunk_3.4}
# Keep only arrays containing less than 500 missing entries
KeepArray= NumberMissingByArray<500
table(KeepArray)
datExpr=datExpr[KeepArray,]
y=y[KeepArray]
#ArrayName[KeepArray]
```


```{r chunk_3.5}
NumberMissingByGene =apply( is.na(data.frame(datExpr)),2, sum)
# One could do a barplot(NumberMissingByGene), but the barplot is empty in this case.
# It may be better to look at the numbers of missing samples using the summary method:
summary(NumberMissingByGene)
# Calculate the variances of the probes and the number of present entries
variancedatExpr=as.vector(apply(as.matrix(datExpr),2,var, na.rm=T))
no.presentdatExpr=as.vector(apply(!is.na(as.matrix(datExpr)),2, sum) )
# Another way of summarizing the number of pressent entries
table(no.presentdatExpr)
# Keep only genes whose variance is non-zero and have at least 4 present entries
KeepGenes= variancedatExpr>0 & no.presentdatExpr>=4
table(KeepGenes)
datExpr=datExpr[, KeepGenes]
#GeneName=GeneName[KeepGenes]
```


```{r chunk_3.6}
# sizeGrWindow(9, 5)
plotClusterTreeSamples(datExpr=datExpr, y=y)
```


```{r}
datExpr[1:3, 1:5]
```


# Standard gene screening illustrates gene selection based on Pearson correlation and shows that the results are not satisfactory


```{r chunk_4.1}
# Load WGCNA package
library(WGCNA)
```


```{r chunk_4.2}
GS1= as.numeric(cor(y, datExpr, use="p"))
# Network terminology: GS1 will be referred to as signed gene significance measure
p.Standard=corPvalueFisher(GS1, nSamples =length(y) )
# since the q-value function has problems with missing data, we use the following trick
p.Standard2=p.Standard
p.Standard2[is.na(p.Standard)]=1
q.Standard=qvalue(p.Standard2)$qvalues
# Form a data frame to hold the results
StandardGeneScreeningResults=data.frame(GeneName,PearsonCorrelation=GS1, p.Standard, q.Standard)
```


```{r chunk_4.3}
NoiseGeneIndicator=is.element( truemodule, c("turquoise", "blue", "yellow", "grey"))+.0
SignalGeneIndicator=1-NoiseGeneIndicator
```


```{r chunk_4.4}
table(q.Standard<.20)
```


```{r chunk_4.5}
mean(NoiseGeneIndicator[q.Standard<=0.20]) 


#=====================================================================================
#
#  Code chunk 6
#
#=====================================================================================


# save.image(file = "Simulated-StandardScreening.RData")
```




# Construction of a weighted gene co-expression network and network modules illustrated step-by-step; includes a discussion of alternate clustering techniques


```{r chunk_5.1}
# Load WGCNA package
library(WGCNA)
# Load additional necessary packages
library(cluster)
```


```{r chunk_5.2}
# here we define the adjacency matrix using soft thresholding with beta=6
ADJ1=abs(cor(datExpr,use="p"))^6
# When you have relatively few genes (<5000) use the following code
k=as.vector(apply(ADJ1,2,sum, na.rm=T))
# When you have a lot of genes use the following code
k=softConnectivity(datE=datExpr,power=6) 
# Plot a histogram of k and a scale free topology plot
#sizeGrWindow(10,5)
par(mfrow=c(1,2))
hist(k)
scaleFreePlot(k, main="Check scale free topology\n")
```


```{r chunk_5.3}
datExpr=datExpr[, rank(-k,ties.method="first" )<=3600]
```


```{r chunk_5.4}
# Turn adjacency into a measure of dissimilarity
dissADJ=1-ADJ1
```


```{r chunk_5.5}
dissTOM=TOMdist(ADJ1)
# collectGarbage()
```


```{r chunk_5.6}
pam4=pam(as.dist(dissADJ), 4)
pam5=pam(as.dist(dissADJ), 5)
pam6=pam(as.dist(dissADJ), 6)
# Cross-tabulte the detected and the true (simulated) module membership:
table(pam4$clustering, truemodule)
table(pam5$clustering, truemodule)
table(pam6$clustering, truemodule)
```


```{r chunk_5.7}
pamTOM4=pam(as.dist(dissTOM), 4)
pamTOM5=pam(as.dist(dissTOM), 5)
pamTOM6=pam(as.dist(dissTOM), 6)
# Cross-tabulte the detected and the true (simulated) module membership:
table(pamTOM4$clustering, truemodule)
table(pamTOM5$clustering, truemodule)
table(pamTOM6$clustering, truemodule)
```


```{r chunk_5.8}
hierADJ=hclust(as.dist(dissADJ), method="average" )
# Plot the resulting clustering tree together with the true color assignment
#sizeGrWindow(10,5);
plotDendroAndColors(hierADJ, colors = data.frame(truemodule), dendroLabels = FALSE, hang = 0.03, 
                    main = "Gene hierarchical clustering dendrogram and simulated module colors" )
```


```{r chunk_5.9}
colorStaticADJ=as.character(cutreeStaticColor(hierADJ, cutHeight=.99, minSize=20))
# Plot the dendrogram with module colors
# sizeGrWindow(10,5);
plotDendroAndColors(hierADJ, colors = data.frame(truemodule, colorStaticADJ),
                    dendroLabels = FALSE, abHeight = 0.99,
                    main = "Gene dendrogram and module colors")
```


```{r chunk_5.10}
branch.number=cutreeDynamic(hierADJ,method="tree")
# This function transforms the branch numbers into colors
colorDynamicADJ=labels2colors(branch.number )
```


```{r chunk_5.11}
colorDynamicHybridADJ=labels2colors(cutreeDynamic(hierADJ,distM= dissADJ, 
                              cutHeight = 0.998, deepSplit=2, pamRespectsDendro = FALSE))

# Plot results of all module detection methods together:
# sizeGrWindow(10,5)
plotDendroAndColors(dendro = hierADJ, 
                    colors=data.frame(truemodule, colorStaticADJ, 
                                     colorDynamicADJ, colorDynamicHybridADJ), 
                    dendroLabels = FALSE, marAll = c(0.2, 8, 2.7, 0.2),
                    main = "Gene dendrogram and module colors")
```


```{r chunk_5.12}
# Calculate the dendrogram
hierTOM = hclust(as.dist(dissTOM),method="average");
# The reader should vary the height cut-off parameter h1 
# (related to the y-axis of dendrogram) in the following
colorStaticTOM = as.character(cutreeStaticColor(hierTOM, cutHeight=.99, minSize=20))
colorDynamicTOM = labels2colors (cutreeDynamic(hierTOM,method="tree"))
colorDynamicHybridTOM = labels2colors(cutreeDynamic(hierTOM, distM= dissTOM , cutHeight = 0.998,
                       deepSplit=2, pamRespectsDendro = FALSE))
# Now we plot the results
# sizeGrWindow(10,5)
plotDendroAndColors(hierTOM, 
               colors=data.frame(truemodule, colorStaticTOM, 
                                 colorDynamicTOM, colorDynamicHybridTOM), 
               dendroLabels = FALSE, marAll = c(1, 8, 3, 1),
               main = "Gene dendrogram and module colors, TOM dissimilarity")
```


```{r chunk_5.13}
tabStaticADJ=table(colorStaticADJ,truemodule)
tabStaticTOM=table(colorStaticTOM,truemodule)
tabDynamicADJ=table(colorDynamicADJ, truemodule)
tabDynamicTOM=table(colorDynamicTOM,truemodule)
tabDynamicHybridADJ =table(colorDynamicHybridADJ,truemodule)
tabDynamicHybridTOM =table(colorDynamicHybridTOM,truemodule)
```


```{r chunk_5.14}
randIndex(tabStaticADJ,adjust=F)
randIndex(tabStaticTOM,adjust=F)
randIndex(tabDynamicADJ,adjust=F)
randIndex(tabDynamicTOM,adjust=F)
randIndex(tabDynamicHybridADJ ,adjust=F)
randIndex(tabDynamicHybridTOM ,adjust=F)
```


```{r chunk_5.15}
colorh1= colorDynamicHybridTOM
# remove the dissimilarities, adjacency matrices etc to free up space
#rm(ADJ1); rm(dissADJ);              
# collectGarbage()
# save.image("Simulated-NetworkConstruction.RData")

```


# Relating modules and module eigengenes to external data illustrates methods for relating modules to external microarray sample traits


```{r chunk_6.1}
# Load WGCNA package
library(WGCNA)
```


```{r chunk_6.2}
datME=moduleEigengenes(datExpr,colorh1)$eigengenes
signif(cor(datME, use="p"), 2)
```


```{r chunk_6.3}
dissimME=(1-t(cor(datME, method="p")))/2
hclustdatME=hclust(as.dist(dissimME), method="average" )
# Plot the eigengene dendrogram
par(mfrow=c(1,1))
plot(hclustdatME, main="Clustering tree based of the module eigengenes")
```


```{r chunk_6.4}
# sizeGrWindow(8,9)
plotMEpairs(datME,y=y)
```


```{r chunk_6.5}
signif(cor(datME, ModuleEigengeneNetwork1[,-1]),2)
```


```{r chunk_6.6}
# sizeGrWindow(8,9)
par(mfrow=c(3,1), mar=c(1, 2, 4, 1))
which.module="turquoise"; 
plotMat(t(scale(datExpr[,colorh1==which.module ]) ),nrgcols=30,rlabels=T,
         clabels=T,rcols=which.module,
         title=which.module )
# for the second (blue) module we use
which.module="blue";  
plotMat(t(scale(datExpr[,colorh1==which.module ]) ),nrgcols=30,rlabels=T,
         clabels=T,rcols=which.module,
         title=which.module )
which.module="brown"; 
plotMat(t(scale(datExpr[,colorh1==which.module ]) ),nrgcols=30,rlabels=T,
         clabels=T,rcols=which.module,
         title=which.module )
```


```{r chunk_6.7}
# sizeGrWindow(8,7);
which.module="green"
ME=datME[, paste("ME",which.module, sep="")]
par(mfrow=c(2,1), mar=c(0.3, 5.5, 3, 2))
plotMat(t(scale(datExpr[,colorh1==which.module ]) ),
         nrgcols=30,rlabels=F,rcols=which.module,
         main=which.module, cex.main=2)
par(mar=c(5, 4.2, 0, 0.7))
barplot(ME, col=which.module, main="", cex.main=2,
        ylab="eigengene expression",xlab="array sample")
```


```{r chunk_6.8}
signif(cor(y,datME, use="p"),2)
```


```{r chunk_6.9}
cor.test(y, datME$MEbrown)
```


```{r chunk_6.10}
p.values = corPvalueStudent(cor(y,datME, use="p"), nSamples = length(y))
```


```{r chunk_6.11}
GS1=as.numeric(cor(y,datExpr, use="p"))
GeneSignificance=abs(GS1)
# Next module significance is defined as average gene significance.
ModuleSignificance=tapply(GeneSignificance, colorh1, mean, na.rm=T)
```


```{r chunk_6.12}
# sizeGrWindow(8,7)
par(mfrow = c(1,1))
plotModuleSignificance(GeneSignificance,colorh1)
```


```{r chunk_6.13}
# collectGarbage()
# save.image("Simulated-RelatingToExt.RData")
```


# Module membership, intramodular connectivity, and screening for intramodular hub genes illustrates using the intramodular connectivity to define measures of module membership and to screen for genes based on network information


```{r chunk_7.1}
# Load WGCNA package
library(WGCNA)
library(cluster)
# Load the previously saved data
# load("Simulated-RelatingToExt.RData");
```


```{r chunk_7.2}
ADJ1=abs(cor(datExpr,use="p"))^6
Alldegrees1=intramodularConnectivity(ADJ1, colorh1)
head(Alldegrees1)
```


```{r chunk_7.3}
colorlevels=unique(colorh1)
# sizeGrWindow(9,6)
par(mfrow=c(2,as.integer(0.5+length(colorlevels)/2)))
par(mar = c(4,5,3,1))
for (i in c(1:length(colorlevels))) 
{
  whichmodule=colorlevels[[i]]; 
  restrict1 = (colorh1==whichmodule);
  verboseScatterplot(Alldegrees1$kWithin[restrict1], 
                     GeneSignificance[restrict1], col=colorh1[restrict1],
                     main=whichmodule, 
                     xlab = "Connectivity", ylab = "Gene Significance", abline = TRUE)
}
```


```{r chunk_7.4}
datKME=signedKME(datExpr, datME, outputColumnName="MM.")
# Display the first few rows of the data frame
head(datKME)
```


```{r chunk_7.5}
FilterGenes= abs(GS1)> .2 & abs(datKME$MM.brown)>.8
table(FilterGenes)
```


```{r chunk_7.6}
dimnames(data.frame(datExpr))[[2]][FilterGenes]
```


```{r chunk_7.7}
# sizeGrWindow(8,6)
par(mfrow=c(2,2))
# We choose 4 modules to plot: turquoise, blue, brown, green. 
# For simplicity we write the code out explicitly for each module.
which.color="turquoise"; 
restrictGenes=colorh1==which.color 
verboseScatterplot(Alldegrees1$kWithin[ restrictGenes], 
                   (datKME[restrictGenes, paste("MM.", which.color, sep="")])^6,
                   col=which.color, 
                   xlab="Intramodular Connectivity", 
                   ylab="(Module Membership)^6")

which.color="blue"; 
restrictGenes=colorh1==which.color
verboseScatterplot(Alldegrees1$kWithin[ restrictGenes],
                   (datKME[restrictGenes, paste("MM.", which.color, sep="")])^6,
                   col=which.color,
                   xlab="Intramodular Connectivity",
                   ylab="(Module Membership)^6")

which.color="brown"; 
restrictGenes=colorh1==which.color
verboseScatterplot(Alldegrees1$kWithin[ restrictGenes],
                   (datKME[restrictGenes, paste("MM.", which.color, sep="")])^6,
                   col=which.color,
                   xlab="Intramodular Connectivity",
                   ylab="(Module Membership)^6")

which.color="green";
restrictGenes=colorh1==which.color 
verboseScatterplot(Alldegrees1$kWithin[ restrictGenes], 
                   (datKME[restrictGenes, paste("MM.", which.color, sep="")])^6,
                   col=which.color, 
                   xlab="Intramodular Connectivity", 
                   ylab="(Module Membership)^6")
```


```{r chunk_7.8}
NS1=networkScreening(y=y, datME=datME, datExpr=datExpr,
         oddPower=3, blockSize=1000, minimumSampleSize=4,
         addMEy=TRUE, removeDiag=FALSE, weightESy=0.5)
```


```{r chunk_7.9}
# network screening analysis
mean(NoiseGeneIndicator[rank(NS1$p.Weighted,ties.method="first")<=100])
# standard analysis based on the correlation p-values (or Student T test)
mean(NoiseGeneIndicator[rank(NS1$p.Standard,ties.method="first")<=100])
```


```{r chunk_7.10}
topNumbers=c(10,20,50,100)
for (i in c(1:length(topNumbers)) ) 
{
  print(paste("Proportion of noise genes in the top", topNumbers[i], "list"))
  WGCNApropNoise=mean(NoiseGeneIndicator[rank(NS1$p.Weighted,ties.method="first")<=topNumbers[i]])
  StandardpropNoise=mean(NoiseGeneIndicator[rank(NS1$p.Standard,ties.method="first")<=topNumbers[i]])
  print(paste("WGCNA, proportion of noise=", WGCNApropNoise, 
        ", Standard, prop. noise=", StandardpropNoise))
  if (WGCNApropNoise< StandardpropNoise) print("WGCNA wins")
  if (WGCNApropNoise==StandardpropNoise) print("both methods tie")
  if (WGCNApropNoise>StandardpropNoise) print("standard screening wins")
}
```


```{r chunk_7.11}
# rm(dissTOM); collectGarbage()
```


```{r chunk_7.12}
#Form a data frame containing standard and network screening results
CorPrediction1=data.frame(GS1,NS1$cor.Weighted)
cor.Weighted=NS1$cor.Weighted
# Plot the comparison
# sizeGrWindow(8, 6)
verboseScatterplot(cor.Weighted, GS1,
         main="Network-based weighted correlation versus Pearson correlation\n",
         col=truemodule, cex.main = 1.2)
abline(0,1)
```


```{r chunk_7.13}
set.seed(2)
nSamples2=2000
MEgreen=rnorm(nSamples2)
scaledy2=MEgreen*ESgreen+sqrt(1-ESgreen^2)*rnorm(nSamples2)
y2=ifelse( scaledy2>median(scaledy2),2,1)
MEturquoise= ESturquoise*scaledy2+sqrt(1-ESturquoise^2)*rnorm(nSamples2)
# we simulate a strong dependence between MEblue and MEturquoise
MEblue= .6*MEturquoise+ sqrt(1-.6^2) *rnorm(nSamples2)
MEbrown= ESbrown*scaledy2+sqrt(1-ESbrown^2)*rnorm(nSamples2)
MEyellow= ESyellow*scaledy2+sqrt(1-ESyellow^2)*rnorm(nSamples2)
# Put together a data frame of eigengenes
ModuleEigengeneNetwork2=data.frame(y=y2,MEturquoise,MEblue,MEbrown,MEgreen, MEyellow)
# Simulate the expression data
dat2=simulateDatExpr5Modules(MEturquoise=ModuleEigengeneNetwork2$MEturquoise,
   MEblue=ModuleEigengeneNetwork2$MEblue,MEbrown=ModuleEigengeneNetwork2$MEbrown,
   MEyellow=ModuleEigengeneNetwork2$MEyellow,
   MEgreen=ModuleEigengeneNetwork2$MEgreen,simulateProportions=simulateProportions1, 
   nGenes=nGenes1)
# recall that this is the signed gene significance in the training data
GS1= as.numeric(cor(y, datExpr, use="p"))
# the following is the signed gene significance in the test data
GS2=as.numeric( cor(ModuleEigengeneNetwork2$y, dat2$datExpr, use="p"))
```


```{r chunk_7.14}
# sizeGrWindow(8,6)
par(mfrow=c(1,1))
verboseScatterplot(GS1,GS2,
       main="Trait-based gene significance in test set vs. training set\n",
       xlab = "Training set gene significance",
       ylab = "Test set gene significance",
       col=truemodule, cex.main = 1.4)
```


```{r chunk_7.15}
EvaluationGeneScreening1 = corPredictionSuccess(
           corPrediction = CorPrediction1, 
           corTestSet=GS2,
           topNumber=seq(from=20, to=500, length=30) )
par(mfrow=c(2,2))
listcomp = EvaluationGeneScreening1$meancorTestSetOverall
matplot(x = listcomp$topNumber,
        y = listcomp[,-1], 
        main="Predicting positive and negative correlations",
        ylab="mean cor, test data", 
        xlab="top number of genes in the training data")
listcomp= EvaluationGeneScreening1$meancorTestSetPositive
matplot(x = listcomp$topNumber,
        y = listcomp[,-1], 
        main="Predicting positive correlations",
        ylab="mean cor, test data", 
        xlab="top number of genes in the training data")
listcomp= EvaluationGeneScreening1$meancorTestSetNegative
matplot(x = listcomp$topNumber,
        y = listcomp[,-1], 
        main = "Predicting negative correlations",
        ylab = "mean cor, test data", 
        xlab = "top number of genes in the training data")
```


```{r chunk_7.16}
relativeCorPredictionSuccess(corPredictionNew = NS1$cor.Weighted,
                             corPredictionStandard = GS1, 
                             corTestSet=GS2,
                             topNumber=c(10,20,50,100,200,500) )
```


```{r chunk_7.17}
# Create a data frame holding the results of gene screening
GeneResultsNetworkScreening=data.frame(GeneName=row.names(NS1), NS1)
# Write the data frame into a file
# write.table(GeneResultsNetworkScreening, file="GeneResultsNetworkScreening.csv", row.names=F,sep=",")
# Output of eigengene information:
datMEy = data.frame(y, datME)
eigengeneSignificance = cor(datMEy, y);
eigengeneSignificance[1,1] = (1+max(eigengeneSignificance[-1, 1]))/2
eigengeneSignificance.pvalue = corPvalueStudent(eigengeneSignificance, nSamples = length(y))
namesME=names(datMEy)
# Form a summary data frame
out1=data.frame(t(data.frame(eigengeneSignificance,
eigengeneSignificance.pvalue, namesME, t(datMEy))))
# Set appropriate row names
dimnames(out1)[[1]][1]="EigengeneSignificance"
dimnames(out1)[[1]][2]="EigengeneSignificancePvalue"
dimnames(out1)[[1]][3]="ModuleEigengeneName"
dimnames(out1)[[1]][-c(1:3)]=dimnames(datExpr)[[1]]
# Write the data frame into a file
# write.table(out1, file="MEResultsNetworkScreening.csv", row.names=TRUE, col.names = TRUE, sep=",")
# Display the first few rows:
head(out1)
```


```{r chunk_7.18}
# Write out gene information
GeneName=dimnames(datExpr)[[2]]
GeneSummary=data.frame(GeneName, truemodule, SignalGeneIndicator,  NS1)
# write.table(GeneSummary, file="GeneSummaryTutorial.csv", row.names=F,sep=",")
# here we output the module eigengenes and trait y without eigengene significances
datTraits=data.frame(ArrayName, datMEy)
dimnames(datTraits)[[2]][2:length(namesME)]=paste("Trait",  
                                             dimnames(datTraits)[[2]][2:length(namesME)], 
                                             sep=".")
# write.table(datTraits, file="TraitsTutorial.csv", row.names=F,sep=",")
rm(datTraits)
# here we output the simulated gene expression data
MicroarrayData=data.frame(GeneName, t(datExpr))
names(MicroarrayData)[-1]=ArrayName
# write.table(MicroarrayData, file="MicroarrayDataTutorial.csv", row.names=F,sep=",")
rm(MicroarrayData)
```


```{r chunk_7.19}
# Perform network screening
NS1GS=networkScreeningGS(datExpr=datExpr, datME = datME, GS=GS1)
# Organize its results for easier plotting
GSprediction1=data.frame(GS1,NS1GS$GS.Weighted)
GS.Weighted=NS1GS$GS.Weighted
# Plot a comparison between standard gene significance and network-weighted gene significance
# sizeGrWindow(8, 6)
par(mfrow=c(1,1))
verboseScatterplot(GS1, GS.Weighted, 
                   main="Weighted gene significance vs. the standard GS\n",
                   col=truemodule)
abline(0,1)
```


```{r chunk_7.20}
EvaluationGeneScreeningGS = corPredictionSuccess(corPrediction=GSprediction1, corTestSet=GS2,
                                    topNumber=seq(from=20, to=500, length=30) )
# sizeGrWindow(8, 6)
par(mfrow=c(2,2))
listcomp= EvaluationGeneScreeningGS$meancorTestSetOverall
matplot(x=listcomp$topNumber,
        y=listcomp[,-1],
        main="Predicting positive and negative correlations",
        ylab="mean cor, test data",
        xlab="top number of genes in the training data")
listcomp= EvaluationGeneScreeningGS$meancorTestSetPositive
matplot(x=listcomp$topNumber, 
        y=listcomp[,-1], 
        main="Predicting positive correlations",
        ylab="mean cor, test data",
        xlab="top number of genes in the training data")
listcomp= EvaluationGeneScreeningGS$meancorTestSetNegative
matplot(x=listcomp$topNumber,
        y=listcomp[,-1],
        main="Predicting negative correlations",
        ylab="mean cor, test data",
        xlab="top number of genes in the training data")
```


```{r chunk_7.21}
# collectGarbage()
# save.image("Simulated-Screening.RData")
```



# Visualization of gene networks
    
    
```{r chunk_8.1}
# Load WGCNA package
library(WGCNA)
library(cluster)
# Load the previously saved data
# load("Simulated-RelatingToExt.RData"); 
# load("Simulated-Screening.RData")
```


```{r chunk_8.2}
cmd1=cmdscale(as.dist(dissTOM),2)
# sizeGrWindow(7, 6)
par(mfrow=c(1,1))
plot(cmd1, col=as.character(colorh1),  main="MDS plot",
     xlab="Scaling Dimension 1", ylab="Scaling Dimension 2")
```


```{r chunk_8.3}
power=6
color1=colorDynamicTOM
restGenes= (color1 != "grey")
diss1=1-TOMsimilarityFromExpr( datExpr[, restGenes], power = 6 )
hier1=hclust(as.dist(diss1), method="average" )
diag(diss1) = NA;
# sizeGrWindow(7,7)
TOMplot(diss1^4, hier1, as.character(color1[restGenes]),
        main = "TOM heatmap plot, module genes" )
```


```{r chunk_8.4}
power=6
color1=colorDynamicTOM
restGenes= (color1 != "grey")
diss1=1-adjacency( datExpr[, restGenes], power = 6 )
hier1=hclust(as.dist(diss1), method="average" )
diag(diss1) = NA;
# sizeGrWindow(7,7)
TOMplot(diss1^4, hier1, as.character(color1[restGenes]),
        main = "Adjacency heatmap plot, module genes" )
```


```{r chunk_8.5}
# sizeGrWindow(7,7)
topList=rank(NS1$p.Weighted,ties.method="first")<=30
gene.names= names(datExpr)[topList]
# The following shows the correlations between the top genes
plotNetworkHeatmap(datExpr, plotGenes = gene.names,
                   networkType="signed", useTOM=FALSE,
                   power=1, main="signed correlations")
```


```{r chunk_8.6}
# sizeGrWindow(7,7)
# The following shows the correlations between the top genes
plotNetworkHeatmap(datExpr, plotGenes = gene.names,
                   networkType="unsigned", useTOM=FALSE,
                   power=1, main="signed correlations")
```


```{r chunk_8.7}
# sizeGrWindow(7,7)
# The following shows the TOM heatmap in a signed network
plotNetworkHeatmap(datExpr, plotGenes = gene.names,
                   networkType="signed", useTOM=TRUE,
                   power=12, main="C. TOM in a signed network")
# The following shows the TOM heatmap in a unsigned network
plotNetworkHeatmap(datExpr, plotGenes = gene.names,
                   networkType="unsigned", useTOM=TRUE,
                   power=6, main="D. TOM in an unsigned network")

```
    



