---
title: "WGCNA step_by_step"
output: 
  html_document:
    toc: true
    code_folding: show
date: "2023-03-31"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.align = "center")
```


## Background

This document is based on "I. Network analysis of liver expression data from female mice: finding modules related to body weight" at the below link.

https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/index.html

It has been adapted to use data generated in this lab for the example as well as other edits where deemed appropriate.


## Load data


```{r}
t1 <- Sys.time()
expData <- read.table(file = "W103first200.tsv", header = TRUE, sep  ="\t")
expData[1:3, 1:8]
#nrow(expData)
#ncol(expData) - 4

datExpr <- expData[, -c(1:4)]
expData <- expData[, 1:4]
#expData[1:3, ]

#names(datExpr) <- paste("Sample", 1:ncol(datExpr), sep  ="")
names(datExpr) <- paste("Sample", sub("^X", "", colnames(datExpr)), sep = "")
rownames(datExpr) <- expData$ID
datExpr <- t(datExpr)
```


Data should be a data.frame where rows are samples and columns are genes.
Row names are samples names and column names are gene names.


```{r}
knitr::kable(datExpr[1:3, 1:6], caption = "**Table 1.**")
```


## Normalization notes

```
colSums(y1);
m1 = apply(y1, 2, median);
hist(y1[,1]);
mm1 = median(m1);
nf0.1 =  m1/mm1;
```

```
nf1 = nf0.1;
yn1 = t(t(y1) / nf1);
apply(yn1, 2, median)
```


## Data overview


```{r, fig.cap="**Figure 1.**", fig.width=12}
# barplot(rowSums(datExpr), las = 3)
barplot(apply(datExpr, MARGIN = 1, median), las = 3)
title( main = "Cumulative expression\nper sample" )
title( ylab = "Expression" )
```


```{r}
sort(rowSums(datExpr))[1:10]
sort(rowSums(datExpr), decreasing = TRUE)[1:10]
```


Dependent variable.


```{r}
phenos <- read.table("Ratios.tsv", header = TRUE, sep = "\t", row.names = 1)
names(phenos) <- paste("Sample", sub("^X", "", colnames(phenos)), sep  ="")
#names(phenos) <- paste("Sample", 1:ncol(phenos), sep  ="")
phenos <- unlist(phenos[1,])
phenos[1:6]
#ncol(phenos)
length(phenos)
hist(phenos, breaks = seq(0, 300, by = 10))
```


```{r, fig.cap="**Figure 2.** No correlation among expression and CBDA:THCA ratio."}
plot( x = datExpr[, 1], y = phenos, xlab = "Expression", ylab = "CBGA:THCA ratio")
cor.test( x = datExpr[, 1], y = phenos)
```


```{r, fig.cap="**Figure 3.** Negative correlation among expression and CBDA:THCA ratio."}
plot( x = datExpr[, 201], y = phenos, xlab = "Expression", ylab = "CBGA:THCA ratio")
cor.test( x = datExpr[, 201], y = phenos)
```



## WGCNA


```{r, message=FALSE}
#library(DESeq2)
library(WGCNA)
#library(dplyr)
#library(readr)
```


## Step-by-step construction of the gene network and identification of modules


```{r}
#plotNetworkHeatmap(datExpr, plotGenes = colnames(datExpr)[1:20])
plotNetworkHeatmap(datExpr, plotGenes = colnames(datExpr))
```



```{r}
# Choose a set of soft-thresholding powers
powers = c(c(1:10), seq(from = 12, to=20, by=2))
# Call the network topology analysis function
enableWGCNAThreads(nThreads = 2)
sft = pickSoftThreshold(datExpr, powerVector = powers, verbose = 5)
# Plot the results:
#sizeGrWindow(9, 5)
par(mfrow = c(1,2))
cex1 = 0.9
# Scale-free topology fit index as a function of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
main = paste("Scale independence"));
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
labels=powers,cex=cex1,col="red");
# this line corresponds to using an R^2 cut-off of h
abline(h=0.90,col="red")
# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
par(mfrow = c(1,1))
```



### The above exploration in powers resulted in '6' and was used below as 'softpower'.


```{r}
softPower = 6
adjacency = adjacency(datExpr, power = softPower)
plot(adjacency)
```



```{r}
# Turn adjacency into topological overlap
TOM = TOMsimilarity(adjacency);
dissTOM = 1-TOM
```


The object `dissTOM` is a gene by gene matrix where the values within the matrix describe a relationship among each pair of genes.
This may require a large amount of time so it may be advantageous to save this result so that future sessions may resume at this point.


```{r}
class(dissTOM)
# write.table(dissTOM, file = "dissTOM.csv", sep = ",", row.names = FALSE, col.names = FALSE)
# dissTOM <- read.csv( file = "dissTOM.csv", header = FALSE )
dissTOM <- read.csv( file = "dissTOM.csv.gz", header = FALSE )
```



### Use distance matrix for `hclust`.

```{r}
# Call the hierarchical clustering function
geneTree = hclust(as.dist(dissTOM), method = "average");
# Plot the resulting clustering tree (dendrogram)
#sizeGrWindow(12,9)
plot(geneTree, xlab="", sub="", main = "Gene clustering on TOM-based dissimilarity",
     labels = FALSE, hang = 0.04)
```


###  Use tree to create 'modules' (i.e., clusters or groups).
Adjust `minModuleSize` to get your desired number of modules.

```{r}
# We like large modules, so we set the minimum module size relatively high:
#minModuleSize = 30
# Or low for small modules.
# Adjust this to get your desired number of modules.
minModuleSize = 3

# Module identification using dynamic tree cut:
dynamicMods = cutreeDynamic(dendro = geneTree, distM = dissTOM,
                            deepSplit = 2, pamRespectsDendro = FALSE,
                            minClusterSize = minModuleSize)
table(dynamicMods)
#dynamicMods[4:10]
#names(dynamicMods[4:10])
#table(names(dynamicMods))
```


```{r}
# Convert numeric lables into colors
dynamicColors = labels2colors(dynamicMods)
table(dynamicColors)
# Plot the dendrogram and colors underneath
# sizeGrWindow(8,6)
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05,
main = "Gene dendrogram and module colors")
```


## Modules of interest

The module 'black' contains our candidate genes.

```{r}

dynamicColors[ grep("W103_g21007.t1", colnames(datExpr)) ]

datExpr[1:3 , dynamicColors == "black" ]

datExpr[1:3 , dynamicColors == "green" ]

dev.off()
plot(datExpr[ , dynamicColors == "green" ][, 1], y = phenos)
cor.test(datExpr[ , dynamicColors == "green" ][, 1], y = phenos)

```



## Eigengenes


```{r}
# Calculate eigengenes
MEList = moduleEigengenes(datExpr, colors = dynamicColors)
MEs = MEList$eigengenes
# Calculate dissimilarity of module eigengenes
MEDiss = 1-cor(MEs)
# Cluster module eigengenes
METree = hclust(as.dist(MEDiss), method = "average")
# Plot the result
#sizeGrWindow(7, 6)
#
plot(METree, main = "Clustering of module eigengenes", xlab = "", sub = "")
#plot(METree)
```



```{r}
MEDissThres = 0.25
# Plot the cut line into the dendrogram
#abline(h=MEDissThres, col = "red")
# Call an automatic merging function
merge = mergeCloseModules(datExpr, dynamicColors, cutHeight = MEDissThres, verbose = 3)
# The merged module colors
mergedColors = merge$colors
# Eigengenes of the new merged modules:
mergedMEs = merge$newMEs
```


```{r}
#sizeGrWindow(12, 9)
#pdf(file = "Plots/geneDendro-3.pdf", wi = 9, he = 6)
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
                    c("Dynamic Tree Cut", "Merged dynamic"),
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05)
#dev.off()
```



```{r}
# Rename to moduleColors
moduleColors = mergedColors
# Construct numerical labels corresponding to the colors
colorOrder = c("grey", standardColors(50))
moduleLabels = match(moduleColors, colorOrder)-1
MEs = mergedMEs
MEs[1:3, ]
# Save module colors and labels for use in subsequent parts
#save(MEs, moduleLabels, moduleColors, geneTree, file = "FemaleLiver-02-networkConstruction-stepByStep.RData")
```




## 3.a Quantifying module–trait associations


```{r}
# Define numbers of genes and samples
nGenes = ncol(datExpr);
nSamples = nrow(datExpr);
# Recalculate MEs with color labels
MEs0 = moduleEigengenes(datExpr, moduleColors)$eigengenes
MEs = orderMEs(MEs0)
#moduleTraitCor = cor(MEs, datTraits, use = "p");

#
moduleTraitCor = cor(MEs, phenos, use = "p");
#moduleTraitCor = cor(MEs, cbind(phenos, phenos), use = "p");

moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples)

tcor <- cbind(moduleTraitCor, 
              moduleTraitPvalue,
              p.adjust(moduleTraitPvalue, method = "BH")
)
colnames(tcor) <- c("Correlation", "P-value", "Adjusted p")
tcor <- tcor[ sort.int(tcor[, "Correlation"], index.return = TRUE, decreasing = TRUE)$ix, ]
tcor
moduleTraitCor <- moduleTraitCor[ sort.int(moduleTraitCor[, 1], index.return = TRUE)$ix, , drop = FALSE]
```


```{r}
MEs[1:3, ]

```





```{r}
#sizeGrWindow(10,6)
# Will display correlations and their p-values
textMatrix = paste(signif(moduleTraitCor, 2), " (", signif(moduleTraitPvalue, 1), ")", sep = "")

dim(textMatrix) = dim(moduleTraitCor)


# png(filename = "pheno_eigengene_heatmap.png", 
#     width = 6.5, height = 5.0, units = 'in',
#     res = 300)
par(mar = c(6, 8.5, 3, 3));
par(mfrow = c(1, 1))
# Display the correlation values within a heatmap plot
labeledHeatmap(Matrix = moduleTraitCor,
  # xLabels = names(datTraits),
  xLabels = "CBGA:THCA",
  #xLabels = c("CBGA:THCA", "CBGA:THCA_2"),
  yLabels = names(MEs),
  ySymbols = names(MEs),
  colorLabels = FALSE,
#  colors = greenWhiteRed(50),
#  colors = viridisLite::magma( n = 100, alpha = 1, begin = 0.1, end = 1),
  colors = c( viridisLite::magma( n = 100, alpha = 1, begin = 0.0, end = 1),
              viridisLite::cividis( n = 100, alpha = 1, begin = 0.0, end = 1, direction = -1)),
  textMatrix = textMatrix,
  setStdMargins = FALSE,
  cex.text = 1.01,
  zlim = c(-1,1),
  main = paste("Module-trait relationships")
  )
#dev.off()
```


```{r}
# Define variable weight containing the weight column of datTrait
# weight = as.data.frame(datTraits$weight_g);
weight = as.data.frame(phenos);
names(weight) = "weight"
# names (colors) of the modules
modNames = substring(names(MEs), 3)
geneModuleMembership = as.data.frame(cor(datExpr, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples))

names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");
geneTraitSignificance = as.data.frame(cor(datExpr, weight, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));
names(geneTraitSignificance) = paste("GS.", names(weight), sep="");
names(GSPvalue) = paste("p.GS.", names(weight), sep="");
```



```{r}
#module = "brown"
#module = "black"
#
module = "green"
column = match(module, modNames);
moduleGenes = moduleColors==module;
#sizeGrWindow(7, 7);
par(mfrow = c(1,1));
verboseScatterplot(abs(geneModuleMembership[moduleGenes, column]),
                   abs(geneTraitSignificance[moduleGenes, 1]),
                   xlab = paste("Module Membership in", module, "module"),
#                   ylab = "Gene significance for body weight",
                   ylab = "Gene significance for phenotype",
                   main = paste("Module membership vs. gene significance\n"),
                   cex.main = 1.2, cex.lab = 1.2, cex.axis = 1.2, col = module)
```






```{r}
colnames(datExpr)[moduleColors=="brown"]
length(colnames(datExpr)[moduleColors=="brown"])
colnames(datExpr)[moduleColors=="black"]

datExpr[, "W103_g288.t1"]

rownames(datExpr)[ datExpr[, "W103_g288.t1"] > 60 ]

MEs[1:3, ]

plot(datExpr[, "W103_g288.t1"], MEs[, "MEblack"])

cor.test(datExpr[, "W103_g288.t1"], MEs[, "MEblack"])

phenos["Sample8.3"]

hist(phenos)
abline( v = phenos["Sample8.3"], col = "#B22222", lwd = 4)

```


Save a module


```{r}
my_module <- datExpr[ , moduleColors=="black"]
my_module[1:3, 1:6]
ncol(my_module)
#write.table(my_module, file = "my_module.csv", sep = ",")
```


```{r}
t99 <- Sys.time()
t99 - t1
```

